The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-10-04 Paper.qmd
Added new dataframe for responses to last child signal and added information from it to the paper.

Child sadness, crying, and temper tantrums were common. Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`. However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`. See @fig-barplot_SignalFreq and @fig-signal_alluvial_plot.

@@

::: {#fig-signal_corrplot}
```{r}
signal_corrplot
```
:::

::: {#fig-signal_alluvial_plot}
```{r}
signal_alluvial_plot
```

@@

We did not predict relationships between child signaling and our demographic control variables, but we did find statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`{r} stats$msadH$EducationLevelYears$str`). See @stbl-main-models s and @fig-effect_edu.

@@

@@

There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (stat). See @fig-effect_last_signal_response and @stbl-last-signal-model.

@@

>>>>

Child sadness, crying, and temper tantrums were common and highly correlated with one another. See @fig-barplot_SignalFreq and @fig-signal_corrplot. Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`. However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`. For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.

@@

::: {#fig-signal_corrplot}
```{r}
signal_corrplot
```
Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
:::

@@

We did not predict relationships between child signaling and our demographic control variables, but we did find statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`{r} stats$msadH$EducationLevelYears$str`). See @stbl-main-models and @fig-effect_edu.

@@

[do we want to mentioned distributions of the actual data?]
Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).

@@

There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (any statistic report here?), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?). See @fig-effect_last_signal_response and @stbl-last-signal-model.

@@

::: {#sfig-signal_alluvial_plot}
```{r}
signal_alluvial_plot
```
